### 2016-04-15 Notes/Notes.tex
misc notes

\documentclass[12pt]{article}

% Emulate MS Word
\usepackage{wordlike}

% One inch margins
\PassOptionsToPackage{margin=1in}{geometry}

% Remove footnote indentation
\usepackage[hang,flushmargin]{footmisc} 
\usepackage{amsmath}

% Add extra blank line between paragraphs, and remove paragraph indentation
%\usepackage[parfill]{parskip}
\parskip = 0\baselineskip

% Header
\usepackage{fancyhdr}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\setlength{\headheight}{57pt}
\lhead{Hannah Director\\HW 5,  Problem 5}

% Add space between text and footnotes section
\setlength{\skip\footins}{0.5in}

% Add a space before the footnote mark
\let\myfootnote\footnote
\renewcommand{\footnote}[1]{\myfootnote{~#1}}

% Don't justify along the right margin
%\raggedright

% Generate placeholder text
\usepackage{lipsum}

\begin{document}

Misc Notes
\begin{itemize}
\item Figure 3: Generate multivariate draws using standard trick (Higdon slide 4, find formal citation)
\item What units are the velocities given in? How did they calculate the complex-valued velocity (observed by instrument directly or calculated based on reversing distances
\item Possible typo: Eq 13  doesn't match Eq 4 (published code matches Eq 4)
\end{itemize} 
\end{document}